---
title: "Multiple Linear Regression"
author: "Jia Ren"
date: "2025-05-09"
output: html_document
---

# Load data

```{r setup, include=FALSE}
df <- read.csv("630finaldata.csv")
df$Strength_picked_r[df$Strength_picked_r %in% 1:5] <- 1
df$Strength_picked_d[df$Strength_picked_d %in% 1:5] <- 1
df$Agility_picked_r[df$Agility_picked_r %in% 1:5] <- 1
df$Agility_picked_d[df$Agility_picked_d %in% 1:5] <- 1
df$Intelligence_picked_r[df$Intelligence_picked_r %in% 1:5] <- 1
df$Intelligence_picked_d[df$Intelligence_picked_d %in% 1:5] <- 1
df$Universal_picked_r[df$Universal_picked_r %in% 1:5] <- 1
df$Universal_picked_d[df$Universal_picked_d %in% 1:5] <- 1
```

# Multiple linear regression

## Simple model

  - Location matters: several regions differ by several minutes in average duration.
	- Hero composition effects: picking Strength or Universal heroes tends to lengthen games (by 30–60s), while Agility picks shorten them.
	- Game dynamics: higher scores substantially extend games (~25–30s per score point).
	- Teamfights: more frequent fights add time.
	- Moderate explanatory power: Adjusted R-squared 0.3023 

```{r}
df$region <- factor(df$region)
df$Strength_picked_r <- factor(df$Strength_picked_r)
df$Strength_picked_d <- factor(df$Strength_picked_d)
df$Agility_picked_r <- factor(df$Agility_picked_r)
df$Agility_picked_d <- factor(df$Agility_picked_d)
df$Intelligence_picked_r <- factor(df$Intelligence_picked_r)
df$Intelligence_picked_d <- factor(df$Intelligence_picked_d)
df$Universal_picked_r <- factor(df$Universal_picked_r)
df$Universal_picked_d <- factor(df$Universal_picked_d)

model <- lm(
  duration ~ region
           + Strength_picked_r + Strength_picked_d
           + Agility_picked_r  + Agility_picked_d
           + Intelligence_picked_r + Intelligence_picked_d
           + Universal_picked_r    + Universal_picked_d
           + first_blood_time
           + dire_score + radiant_score
           + exp_15min
           + teamfight_duration
           + Tteamfight_deaths
           + teamfight_frequency,
  data = df
)

summary(model)

par(mfrow = c(2,2))
plot(model)
```

## Improved models

  - exp_15min shows U shape (See Marginal Linear Regression file): use polynomial terms
    poly(exp_15min, 2)2  becomes strongly negative. The match duration peaks at some mid‑range of exp_15min and then declines at the extremes.
  - Add interaction terms `dire_score*radiant_score`: Dire score and radiant score are in competing relation
  - Add interaction terms `teamfight_duration*teamfight_frequency`: More frequent team fights may add to the total duration of team fights.
  - The model provides better fit: higher Adjusted R-squared (0.3722), lower RSE. 
  
  
```{r}
df$region <- factor(df$region)
df$Strength_picked_r <- factor(df$Strength_picked_r)
df$Strength_picked_d <- factor(df$Strength_picked_d)
df$Agility_picked_r <- factor(df$Agility_picked_r)
df$Agility_picked_d <- factor(df$Agility_picked_d)
df$Intelligence_picked_r <- factor(df$Intelligence_picked_r)
df$Intelligence_picked_d <- factor(df$Intelligence_picked_d)
df$Universal_picked_r <- factor(df$Universal_picked_r)
df$Universal_picked_d <- factor(df$Universal_picked_d)

model_2 <- lm(
  duration ~ region
           + Strength_picked_r + Strength_picked_d
           + Agility_picked_r  + Agility_picked_d
           + Intelligence_picked_r + Intelligence_picked_d
           + Universal_picked_r    + Universal_picked_d
           + first_blood_time
           + dire_score + radiant_score + dire_score * radiant_score
           + poly(exp_15min, 2)
           + teamfight_duration
           + Tteamfight_deaths
           + teamfight_frequency + teamfight_duration * teamfight_frequency,
  data = df
)

summary(model_2)

par(mfrow = c(2,2))
plot(model_2)
```
  